---
title: "FIES scale for analysis of Food Insecurity using the Rasch Model"
output: html_notebook
---

## Importing the Libraries
```{r}
library("RM.weights")
```



## Inserting the Datasets
#### The dataset includes the FIES data (Food Insecurity Experience based Scale), sampling weights, and some demographic variables for a sample country (Country1). Data have been collected by the Gallup World Poll.

```{r}
# Country 1 data
load("C:/Users/Asus/Desktop/Final_FIES/data.FAO_country1.rda")

# Country 2 data
load("C:/Users/Asus/Desktop/Final_FIES/data.FAO_country2.rda")

# Country 3 data
load("C:/Users/Asus/Desktop/Final_FIES/data.FAO_country3.rda")

# Country 4 data
load("C:/Users/Asus/Desktop/Final_FIES/data.FAO_country4.rda")
```

# Analysing Country 1
## Questionnaire data and weights
```{r}
# Matrix of questionnaire data and weights (for the Rasch analysis)
XX.country1 = data.FAO_country1[,1:8]
wt.country1 = data.FAO_country1$wt
```

## Creating the Rasch Model for Country 1
```{r}
# Fit weighted Rasch
rr.country1 = RM.w(XX.country1, wt.country1)

# Fit unweighted Rasch
rr.country1.nw = RM.w(XX.country1)
```

### Displaying important parameters

```{r}
# Item severity
rr.country1$b
```

```{r}
# Item standard error
rr.country1$se.b
```

```{r}
# Respondent severity for each raw score
rr.country1$a
```

```{r}
# Respondent measurement error for each raw score
rr.country1$se.a
```

```{r}
# Item infit
rr.country1$infit
```

```{r}
# Item outfit
rr.country1$outfit
```

```{r}
# Rasch reliability based on observed distribution of cases across raw scores
rr.country1$reliab
```

```{r}
# Rasch reliability based on equal proportion of cases in each
# non-extreme raw score (more comparable across datasets)
rr.country1$reliab.fl
```

```{r}
# Checking conditional independence: residual correlation matrix
rr.country1$res.cor
```

## Equating against 2014-2016 FAO global standard for the Food Insecurity Experience scale.

#### Iterative Equating.
```{r}
# Iterative equating
ee1=equating.fun(rr.country1, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE)

# Common and Unique Items
ee1$common
```

#### Non-Iterative Equating.
```{r}
ee2=equating.fun(rr.country1, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE, iterative = FALSE,excl.prior1=1,excl.prior2=1)

# Common and Unique Items
ee2$common
```

## Item Characteristic Curve (ICC)

```{r}
# Taking the Item standard error.
b=rr.country1$b

# ICC values
icc = ICC.fun(b)$icc

# ICC plot
icc = ICC.fun(b, plot = TRUE)
```

## Displaying prevelance of FOOD Insecurity for Country 1.
### Prevalence rate (Mod+Sev)	& Prevalence rate (Sev) are displayed below
### Fmod+sev and Fsev


```{r}
## Prevalence rates
# Prevelance rates after Iterative Equating
ee1$prevs

# Prevelance rates after Non-Iterative Equating
ee2$prevs
```

=====================================================================================================
=====================================================================================================

# Analysing Country 2
## Questionnaire data and weights
```{r}
# Matrix oftionnaire data and weights (for the Rasch analysis)
XX.country2 = data.FAO_country2[,1:8]
wt.country2 = data.FAO_country2$wt
```

## Creating the Rasch Model for Country 2
```{r}
# Fit weighted Rasch
rr.country2 = RM.w(XX.country2, wt.country2)

# Fit unweighted Rasch
rr.country2.nw = RM.w(XX.country2) 
```

### Displaying important parameters

```{r}
# Item severity
rr.country2$b
```

```{r}
# Item standard error
rr.country2$se.b
```

```{r}
# Respondent severity for each raw score
rr.country2$a
```

```{r}
# Respondent measurement error for each raw score
rr.country2$se.a
```

```{r}
# Item infit
rr.country2$infit
```

```{r}
# Item outfit
rr.country2$outfit
```

```{r}
# Rasch reliability based on observed distribution of cases across raw scores
rr.country2$reliab
```

```{r}
# Rasch reliability based on equal proportion of cases in each
# non-extreme raw score (more comparable across datasets)
rr.country2$reliab.fl
```

```{r}
# Checking conditional independence: residual correlation matrix
rr.country2$res.cor
```

## Equating against 2014-2016 FAO global standard for the Food Insecurity Experience scale.

#### Iterative Equating.
```{r}
# Iterative equating
ee1.country2=equating.fun(rr.country2, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE)

# Common and Unique Items
ee1.country2$common
```

#### Non-Iterative Equating.
```{r}
ee2.country2=equating.fun(rr.country2, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE, iterative = FALSE,excl.prior1=3,excl.prior2=3)

# Common and Unique Items
ee2.country2$common
```

## Item Characteristic Curve (ICC)

```{r}
# Taking the Item standard error.
b=rr.country2$b

# ICC values
icc.country2 = ICC.fun(b)$icc

# ICC plot
icc.country2 = ICC.fun(b, plot = TRUE)
```

## Displaying prevelance of FOOD Insecurity for Country 2.
### Prevalence rate (Mod+Sev)	& Prevalence rate (Sev) are displayed below
### Fmod+sev and Fsev


```{r}
## Prevalence rates
# Prevelance rates after Iterative Equating
ee1.country2$prevs

# Prevelance rates after Non-Iterative Equating
ee2.country2$prevs
```

=====================================================================================================
=====================================================================================================

# Analysing Country 3
## Questionnaire data and weights
```{r}
# Matrix oftionnaire data and weights (for the Rasch analysis)
XX.country3 = data.FAO_country3[,1:8]
wt.country3 = data.FAO_country3$wt
```

## Creating the Rasch Model for Country 3
```{r}
# Fit weighted Rasch
rr.country3 = RM.w(XX.country3, wt.country3)

# Fit unweighted Rasch
rr.country3.nw = RM.w(XX.country3) 
```

### Displaying important parameters

```{r}
# Item severity
rr.country3$b
```

```{r}
# Item standard error
rr.country3$se.b
```

```{r}
# Respondent severity for each raw score
rr.country3$a
```

```{r}
# Respondent measurement error for each raw score
rr.country3$se.a
```

```{r}
# Item infit
rr.country3$infit
```

```{r}
# Item outfit
rr.country3$outfit
```

```{r}
# Rasch reliability based on observed distribution of cases across raw scores
rr.country3$reliab
```

```{r}
# Rasch reliability based on equal proportion of cases in each
# non-extreme raw score (more comparable across datasets)
rr.country3$reliab.fl
```

```{r}
# Checking conditional independence: residual correlation matrix
rr.country3$res.cor
```

## Equating against 2014-2016 FAO global standard for the Food Insecurity Experience scale.

#### Iterative Equating.
```{r}
# Iterative equating
ee1.country3=equating.fun(rr.country3, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE)

# Common and Unique Items
ee1.country3$common
```

#### Non-Iterative Equating.
```{r}
ee2.country3=equating.fun(rr.country3, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE, iterative = FALSE,excl.prior1=7,excl.prior2=7)

# Common and Unique Items
ee2.country3$common
```

## Item Characteristic Curve (ICC)

```{r}
# Taking the Item standard error.
b=rr.country3$b

# ICC values
icc.country3 = ICC.fun(b)$icc

# ICC plot
icc.country3 = ICC.fun(b, plot = TRUE)
```

## Displaying prevelance of FOOD Insecurity for Country 3.
### Prevalence rate (Mod+Sev)	& Prevalence rate (Sev) are displayed below
### Fmod+sev and Fsev


```{r}
## Prevalence rates
# Prevelance rates after Iterative Equating
ee1.country3$prevs

# Prevelance rates after Non-Iterative Equating
ee2.country3$prevs
```

=====================================================================================================
=====================================================================================================


# Analysing Country 4
## Questionnaire data and weights
```{r}
# Matrix oftionnaire data and weights (for the Rasch analysis)
XX.country4 = data.FAO_country4[,1:8]
wt.country4 = data.FAO_country4$wt
```

## Creating the Rasch Model for Country 4
```{r}
# Fit weighted Rasch
rr.country4 = RM.w(XX.country4, wt.country4)

# Fit unweighted Rasch
rr.country4.nw = RM.w(XX.country4) 
```

### Displaying important parameters

```{r}
# Item severity
rr.country4$b
```

```{r}
# Item standard error
rr.country4$se.b
```

```{r}
# Respondent severity for each raw score
rr.country4$a
```

```{r}
# Respondent measurement error for each raw score
rr.country4$se.a
```

```{r}
# Item infit
rr.country4$infit
```

```{r}
# Item outfit
rr.country4$outfit
```

```{r}
# Rasch reliability based on observed distribution of cases across raw scores
rr.country4$reliab
```

```{r}
# Rasch reliability based on equal proportion of cases in each
# non-extreme raw score (more comparable across datasets)
rr.country4$reliab.fl
```

```{r}
# Checking conditional independence: residual correlation matrix
rr.country4$res.cor
```

## Equating against 2014-2016 FAO global standard for the Food Insecurity Experience scale.

#### Iterative Equating.
```{r}
# Iterative equating
ee1.country4=equating.fun(rr.country4, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE)

# Common and Unique Items
ee1.country4$common
```

#### Non-Iterative Equating.
```{r}
ee2.country4=equating.fun(rr.country4, spec.com1=c(1:8),spec.com2=c(1:8), write.file=TRUE, plot=TRUE, iterative = FALSE,excl.prior1=1,excl.prior2=1)

# Common and Unique Items
ee2.country4$common
```

## Item Characteristic Curve (ICC)

```{r}
# Taking the Item standard error.
b=rr.country4$b

# ICC values
icc.country4 = ICC.fun(b)$icc

# ICC plot
icc.country4 = ICC.fun(b, plot = TRUE)
```

## Displaying prevelance of FOOD Insecurity for Country 4.
### Prevalence rate (Mod+Sev)	& Prevalence rate (Sev) are displayed below
### Fmod+sev and Fsev


```{r}
## Prevalence rates
# Prevelance rates after Iterative Equating
ee1.country4$prevs

# Prevelance rates after Non-Iterative Equating
ee2.country4$prevs
```


























